Dynamic capabilities and competitive advantage

### 2019-11-15 article4_jbr.rnw
Started writing up the results

\documentclass[review,fleqn]{elsarticle}
\usepackage[textsize=tiny,colorinlistoftodos]{todonotes}
\usepackage{subfig}
\usepackage{subcaption}

\input{preamble.tex}
%%\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\setlength{\mathindent}{0pt} %% NB MAYBEREMOVE BEFORE SUBMISSION

@@

%\address[mymainaddress]{1600 John F Kennedy Boulevard, Philadelphia}
%\address[mysecondaryaddress]{360 Park Avenue South, New York}

@@

%% Extant literature on dynamic capabilities is battling with a fundamental conception and
%% prediction of the theory, namely if, when and how dynamic capabilities can lead to competitive
%% advantage. 
%% This paper is about how apsobritve capacity can come about as strategic posture
%https://arrow.dit.ie/cgi/viewcontent.cgi?article=1009&context=buschmancon

@@

The conventional mechanism explored in the extant literature is related to changes in the
firm`s resource base \citep{Eriksson2014,Schilke2018}. This mechanism is also included in
the dominating definition of dynamic capabilities found in \citep{Helfat2007}: 'A dynamic
capability is the capacity of an organization to purposefully create, extend or modify its
resource base' (p. 4). And although this definition opens up for a variety of components
included in the term 'resource base', a point we discuss later in this paper, most of the
subsequent contributions have focused on changes in operational routines (ref),
capabilities (ref), practices (ref) or tangible assets (ref). This focus, albeit very
important to the development of dynamic capabilities literature, seems to leave out a
considerable part of a firm`s 'resource base', namely those resources that are tied to how
the employees of the firm sees their mission, internalize information and effectuates
change. These traits of the organization is very much related to concepts of
'organizational culture' (ref), 'employee cognition' (ref)' 'organizational genes'
\citep{Nelson1982} and 'sense-making' \citep{Weick1995} to name but a few. However, in
relations with dynamic capabilities one promising avenue for exploring mechanisms beyond
the conventional ones described above, makes a clear distinction between \emph{behavioral}
and \emph{non-behavioral} objects of change \citep{Verona2011}. These non-behavioral
objects of change contain psychological objects such as cognitive frames, motivation and
identity that are important for strategy formation and implementation (Gavetti etc, vince
and gabriel ch 15 of book by verona) as they 'concerns how decisions or actions are
shaped, subverted and transformed by emotions, and (..) determine the "way we do things
here"' (Vince and gabriel). In this respect they work as firm resources in generating
outcomes such as performance and competitive advantage. This notion is perhaps expressed
most clearly by Peter Drucker`s famous quote about how 'culture eats strategy for lunch'
(ref) insofar as . Moreover, these non-behavioral objects of change are themselves far
from static and are found to change over time as responses to external change
\citep{Zollo2016} thus making them a natural part of the mechanisms underpinning how
dynamic capabilities work \citep{Schilke2018,Verona2011}. Still, far less is know about
the non-behavioral objects of change and how they work as a mechanism in the dynamic
capabilities theory.

@@

To shed light on this important question we theorize on the role of a particular type of
non-behavioral object of change, \emph{organizational cognition}, and how it acts as a
mediator between dynamic capabilities and firm competitive advantage and hence has a
pivotal place in the literature. We build on important work on the concept of
organizational cognition (ref), non-behavioral object of change (ref) and blablabla (ref),
and hypothesize about how organizational cognition acts as a mediator in the same way as
conventional behavioral objects of change (such as routines and capabilities) and that
these two are complementary to a certain degree. Moreover, we also suggest how
organizational cognition could also influence routines and capabilites directly with
further implications for competitive advantage.

@@

@@

DC as resource orchestration - influencing capabilities
DC as cognitive programmer - inluencing cognitive frames

@@

Satya Nadella is very much on the same page as recent contributions on how cognitive
frames of an organization (i.e. its non-behavioral object of action) is a pivotal part in
how firm resources shape outcomes such as performance and competitive advantage
\citep{Barney1991a,Barney2001}, and how it consequently plays a role in how dynamic
capabilities work along these lines \cite{Verona2011}. Microsoft under Nadella did
transform the company by changing the behavioral objects of action (i.e. capabilities and
routines), but also through changing the non-behavioral objects of action (i.e. the
cognitive frames). In other words, the story of Microsoft`s transformation suggest that
dynamic capabilities worked through mediating factors to influence competetive advantage,
but that thes mediating factors exist as two distinctly different concepts working in
tandem. Figure \ref{fig:fig1} presents a conceptual depiction of the paths running through
these two concepts of \emph{behavioral} and \emph{cognitive} objects of action
\citep{Zollo2016}. We use a familiar term for the conventional understanding of dynamic
capabilities mechanism as 'resource orchestration' (ref) as a collective term to describe
the behavioral objects of action (ref). This idea of orchestrating resources are
particularly illuminating in that it sets an organization`s dynamic capabilities as a
director conducting new outcomes by orchestrating the use of firm resources, and it is in
line with a range of the most common definitions of dynamic capabilities
\citep{Schilke2018}. The non-behavioral objects of action has no similar agreed upon
analogy, so for the purpose of our analysis we envision this mechanism to work more as a
cognitive process enabling the organization to tear away from their current understanding
and ways in which things are done (Vince and Gabriel). Whereas the analogy of resource
orchestration has its counterpart in the making of art (i.e. music, movies or theatrical
plays), the non-behavioral objects of action are more aligned to psychology and cognitive
therapy. The process is all about changing an organization`s thinking from the automated
and well-known, to the more cognitively aware, what Kahneman and Tversky coins 'system two'
thinking (ref). This change of cognitive frames will thus be more similar to a
reprogramming effort and we henceforth refer to the process of changing the non-behavioral
objects of action (e.g. cognitive frames, culture, etc) as 'cognitive programming'.

@@

RESOURCE ORCH is HITT2011 among other
\subsection{Cognitive programming}
Juxtaposed to the experiences from Microsoft, the case of Polaroid brings another set of
experiences to our discussion on non-behavioral objects of action (see \citep{Tripsas2000}
for a detailed account). Cognitive inertia on the part of the organization, not the top
management, \todo{Check this - maybe remove the sentence on top management} played a
pivotal part in the failing of the company`s digital imaging effort
\citep{Verona2011}. Similar patterns are found in Kodak and Anderson Consulting (ref
Hederson and Kaplan 2005), and Smith Corona (Danneels 2010). These examples highlight the
absence of 'dynamic capabilities specific to the adaptation of cognitive frames, first,
and consequently of operating processes' \citep[p. 541]{Verona2011}.

Looking at dynamic capabilities through the functions of sensing, seizing and
transformation \cite{Teece2007} makes an obvious case for resource orchestration which is also highlighted
in the extant literature. Moreover, as discussed above, these functions are also requiring
a certain non-behavioral object of action in order to be fully efficient. To achieve a
competitive edge firms will have to be able to sense and seize, which require a certain
organizational cognitive frame to be efficient. 'Dynamic capabilities are about doing the
right things, at the right time, based on new product (and process) development, unique
managerial orchestration processes, a strong and {\bf change-oriented organizational culture}'
(our emphasis) \cite[p]{Teece2014}.

To see how dynamic capabilities acts through cognitive objects of action (i.e. as a
cognitive programmer) we consider two distinctly different functions of dynamic
capabilities related to their microfoundations \citep{Teece2007}. the first function is
related to the discovering and capturing opportunities (e.g. the ability of the
organization to sense and seize opportunities (ibid)) which is very much related to the
concept of exploration \citep{March1991}. Organizations with high levels of dynamic
capabilities geared towards sensing and seizing opportunities are doing so through
\emph{resource orchestration}, e.g. by creating, extending and modifying the resource
base, as discussed extensively in the extant literature
\citep{Helfat2007,DiStefano2014,Protogerou2012,Schilke2018}. However, sensing and seizing
opportunities in the market relates to discovering market disequilibrium
\citep{Kirzner1997} as well as upsetting existing equilibria \citep{Schumpeter1934}. The
process of sensing and seizing 'involves understanding latent demand, the structural evolution of
industries and markets, and likely supplier and competitor responses'
\citep[p. 1322]{Teece2007} as well as . Individuals are the unit of analysis embedding these abilities
and if left up to these individuals without a higher organizing framework, the firm will
be very vulnerable because of knowledge asymmetries and incentive problems (alchian and
demsetz 1972) and lead to sub optimal rent-appropriation on the part of the individual
employee \citep{Blyler2003}. A similar argument applies to the function of \emph{seizing}
where the organization is set to mobilize resources for the purpose of gaining from these
opportunities. In order for value to be captured from opportunities, a string of
conditions have to be met \citep{Teece2017}, but the evolution of a 'strong and
change-oriented organizational culture' \citep[p. 331]{Teece2014} is of pivotal
importance. Alas the sensing process requires new knowledge from a wide range of sources
in the organization a lot of this knowledge is tacit and can lead to rent-appropriation
from information asymmetries \citep{Blyler2003} if not sufficiently aligned with the
cognitive frames of the organization.

One consequence is a ‘program persistence bias.’ Its corollary is various forms of
‘anti-innovation bias,’ including the ‘anti- cannibalization’ basis discussed in a later
section. Program persistence refers to the funding of pro- grams beyond what can be
sustained on the merits,

However, in this endeavour
simply adatping resources is a necessary but not sufficient condition for obtaining the
needed strategic change.

The functions of dynamic capabilities as highlighted above is thoroughly discussed in
extant literature (refs), but a quick recap is useful.

DC geared towards creating new or reconfiguring old (exploration vs exploitation) but true
competetiveness over time stems from ambidexterity- link to research on cultural ambidecterity

The concept of sensing and seizing has analogous meaning in theoretical traditions tangent
to DC. Notably entrepreneurship \citep{Hitt2011} and organizational theory
\citep{March1991}. \cite{Hitt2011} states that "entrepreneurship involves identifying and
exploiting opportunities" (ibid p 59) where identification is tied to the process of
discovery \citep{Shane2000}. Based on the definitions of opportunity discovery as DEF HERE
juxtaposed to the definition of sensing as "a scanning, creation, learning, and
interpretive activity" \citet[p. 1322]{Teece2007} we argue that they are similar in nature
and usable for exploring the importance of the latter in the innovation process. It also
bears resemblance to the concept of organizational exploration as searching, risk-taking,
discovery, experimentation, prototyping, and flexibility \citep{March1991}. Independent of
the nature of opportunities to be sensed as being either objective
\citep{Alvarez2007,Shane2000} or created \citep{Alvarez2007}, the process of sensing does
require resources and capabilities for search (ref) and a clear affinity between
opportunity and a project exist (casson and Wadeson in Mcmullen2007) suggesting s clear
inter dependency between routines and resources, and higher order capabilities of sensing
\citep{Teece2007} and making sense of the opportunity
\citep{Weick1995,March1991}. Consequently, we would expect the sensing class of dynamic
capabilities to be positively related to innovation on the one hand, and reinforcing the
impact from innovation input simultaneously.\newline

To see how dynamic capabilities work to adapt cognitive frames in addition to operational
capabilities to influence competetive advantage, it is useful to turn to their
microfoundations \citep{Teece2007}. In this framework dynamic capabilities work through
three expressions. First, sensing seizing. more here..

Each of these dimensions relates to a function of cognitive frames within the firm beyond
the cognitive frames of the dynamic capability itself as explored by previous
contributions in general \citep{Teece2007} and towards management specifically
\citep{Helfat2015,Helfat2015a}. First, sensing opportunities in the market relates to
discovering market disequilibrium \citep{Kirzner1997} as well as upsetting existing
equilibria \citep{Schumpeter1934}. This process is very much at the core of dynamic
capabilities as a function of management, but its function calls for a very close
relationship with the rest of the organization both in terms of resource orchestration and
cognitive programming. The process of sensing 'involves understanding latent demand, the
structural evolution of industries and markets, and likely supplier and competitor
responses' \citep[p. 1322]{Teece2007}. Individuals are the unit of analysis embedding
these abilities and if left up to these individuals without a higher organizing
framework, the firm will be very vulnerable because of knowledge asymmetries and incentive
problems (alchian and demsetz 1972) and lead to sub optimal rent-appropriation on the part
of the individual employee \citep{Blyler2003}.

A similar argument applies to the function of \emph{seizing} where the organization is set
to mobilize resources for the purpose of gaining from these opportunities. In order for
value to be captured from opportunities, a string of conditions have to be met
\citep{Teece2017}, but the evolution of a 'strong and change-oriented organizational
culture' \citep[p. 331]{Teece2014} is of pivotal importance. Alas the sensing process
requires new knowledge from a wide range of sources in the organization a lot of this
knowledge is tacit and can lead to rent-appropriation from information asymmetries
\citep{Blyler2003} if not sufficiently aligned with the cognitive frames of the
organization.

Finally, the \emph{transformation} function of dynamic capabilities are related to the
idea of continuously renewal of the resource base in order to stay competitive. This continuous
process requires the development and maintenance of a share cognitive frame such as a
culture, which is typically found in production processes such as LEAN and Six Sigma
(ref).

Looking at dynamic capabilities through the functions of sensing, seizing and
transformation makes an obvious case for resource orchestration which is also highlighted
in the extant literature. Moreover, as discussed above, these functions are also requiring
a certain non-behavioral object of action in order to be fully efficient. To achieve a
competitive edge firms will have to be able to sense and seize, which require a certain
organizational cognitive frame to be efficient. 'Dynamic capabilities are about doing the
right things, at the right time, based on new product (and process) development, unique
managerial orchestration processes, a strong and change-oriented organizational culture'
\cite[p]{Teece2014}.

It is argued that contextual ambidexterity is grounded in the type of organizational culture (Ghoshal and Bartlett 1994; Gibson and Birkinshaw 2004; Simsek et al. 2009) that promotes both creativity and discipline (Jelinek and Schoonhoven 1993), or both the presence of different knowledge and the integration of multiple perspectives to develop a cohesive point of view (Eisenhardt and Schoonhoven 1990).

ambidextrous organizational culture wang rafiq 2014

The enterprise will be vulner- able if the sensing, creative, and learning functions are left to the cognitive traits of a few individuals

Three types of managerial activities can make a capability dynamic: sensing (which means
identifying and assessing opportunities outside your company), seizing (mobilizing your
resources to capture value from those opportunities), and transforming (continuous
renewal).

Whereas ordinary capabilities are about doing
things right, dynamic capabilities are about doing the right things, at the right time, based on new product (and process) development, unique mana- gerial orchestration processes, a strong and change- oriented organizational culture, and a prescient assessment of the business environment and tech- nological opportunities.
teece 2014

The strength of a firm's dynamic capabilities determines the speed and degree (and
associated cost) of aligning the firm's resources—including its business model(s)—with
customer needs and aspirations. To achieve this, organizations must be able to
continuously sense and seize opportunities, and to periodically transform aspects of the
organization and culture so as to be able to proactively reposition to address yet newer
threats and opportunities as they arise. Terse 2018

Dynamic capabilities are hard for rivals to replicate because they are built on the
idiosyncratic characteristics of entrepreneurial managers and the history-honed routines
and culture of the organization (Teece, 2014a). In addition, there is the uncertain
imitability of a complex system that even those directly involved may not fully understand
(Lippman and Rumelt, 1982). Because they are a unique and valuable general-purpose
resource, strong dynamic capabilities can serve as a firm foundation for sustainable
competitive advantage. This is especially true the more deeply embedded the capabilities
are in the organization, and the less they are resident only in the top management
team. Teece 2018

 Managers must articulate a vision and establish an appropriate organizational culture and an incentive system that will promote organizational identification and loyalty (Augier and Teece, 2009).

Put differ- ently, selecting the right ‘architecture’ for a busi- ness requires not just understanding the choices available; it also requires assembling the evidence needed to validate conjectures and hunches about costs, customers, competitors, complementors, dis- tributors, and suppliers. Designing

Distinct rolle

H1: Cognitive frames mediates the effect of dynamic capabilities on competetive advantage
above and beyond mediating effect of operational capabilities

H2: Cognitive frames has a direct impact on operational capabilities

H3: Something to do with dynamism - importance of cognitive frames increases in dynamism

Minor hypotheses: On

The role of cognitive frames in strategy: Schilling 2018 abstraction

Proactiveness
Coff and Blyler - who reaps the reward from rents). cognitive frames common -
proactiveness - culture can prohibit the rent appropriation happening

Cognitive programming as rent appropriation inihibitor

Cognitive programming as implementation catalyst (complement to OC)

Cognitive programming as alignment mechanism (link to OC)

THIS IS HOW FAR I CAME
\newpage

Ethnographic study on cognitive farmes - still top management team
http://knowledge.wharton.upenn.edu/wp-content/uploads/2013/09/13481.pdf

Implications: Maybe proactiveness cog frames can help in the paradox schreyogg, and the
elephant.

We suggest that a systematic socio-cognitive approach to “competitive sensemaking” has
been absent from theory and research on competitive strategy. We define competitive
sensemaking as the social and cognitive processes that underlie how firms detect, define,
and conceptualize their competitive relationships with other firms. Competitive
sensemaking is a subset of the more general process of strategic sensemaking, which is the
“making plausible sense” of the broad array of stimuli and circumstances that characterize
complex market situations. Using the value-based view of value creation and capture as a
conceptual base for our arguments, we unpack four cognitive underpinnings of competitive
sensemaking: mental time travel, comparability, counterfactual reasoning, and stories. We
then show how these four components were differentially involved in shaping competitive
sensemaking in four actual market situations. In doing so, we illustrate how competitive
sensemaking provides fundamental inputs into the value creation and value capture
process. We conclude the paper by drawing out the implications of competitive sensemaking
for strategy theory and research.

Some strategic decisions will long be remembered for being remarkably successful, providing dramatic benefits to customers, investors, or social welfare. The individuals who make these decisions are often called "visionary." But what enables some people to be visionary? Is it a gift or skill, and can it be learned? In this paper I identify three main cognitive processes that underlie the creation of visionary strategy: abstraction, idealism, and long paths of analytical reasoning. Using a combination of examples and cognitive science, I explain how they work individually and in combination, and how they may be cultivated.
Schiller

Cognition explain organizational inertia garud and rappa
Sharing of dominant logic pralahad and bettis (actually managmenet but maybe org as well)
Tripsas and Gavetti 2000 focus on management cognition in digital imaging

power of analogical reasoning prevailing in teh strategy makin in novel and complex worlds

(Gavetti, 2012; Gavetti et al., 2005; Tripsas and Gavetti, 2000), awaiting from LB

DCs are viewed as stable patterns of action that poten- tially build a string of temporary
advantages by adding, integrating, and reconfi guring resources, which amount to sustained
advantage once the entire pattern is taken into account (Blyler and Coff, 2003)

Again, some acquiring fi rms (not many) undergo these motivational change processes (Gottschalg and Zollo, 2007) by

would for the larger part argue that 
including the 33000 people at nokia / cultural change rr

Capabilities changes reinvent productivity and business processs, as well as introcuing
new capabiltieis. 
Cloud based capabilities and infrastructure

'AS I conlcluded my talk that morning in orlando, i focused on what would be our grandest
endeavor / the biggest hurdle, transforming the microsofg culture. p 89

Defined culture as a 'complex system of individual mindsets' culture is how an
organization thinks and acts, but indiciduals shape it. p 90

%https://news.microsoft.com/2014/02/04/satya-nadella-email-to-employees-on-first-day-as-ceo/
%https://qz.com/work/1539071/how-microsoft-ceo-satya-nadella-rebuilt-the-company-culture/

%{\bf How is a direct link between dynamic capabiliites and competetive advantage working
%  and under which conditions does it come into play or not?}

In the face of a growing consensus on the definition (ref) and outcome (ref) of dynamic
capabilities as working through mediators, the presence of direct effects argued by others
(ref) seems puzzling. However, the broader empirical literature on the antecedents of firm
performance offers an interesting point of departure. This literature concerns itself with
the relative importance of factors at the firm-level on the one hand, and those at the
industry level at the other.

The debate over the relative role of different levels of factors in accounting for the
major share of the variance in firms performance has been of particular importance in the
field of strategic management, most notably in the Strategic Management Journal
(M. A. Fitza 2014; Hansen and Wernerfelt 1989; McGahan and Porter 1997; Rumelt
1991). Overall, the accumulated empirical evidence show that firm factors matter far more
for firm performance than industry factors (McGahan and Porter 2002). These stylized facts
have been used as empirical support for the importance of idiosyncratic firm resource and
capabilities; leading up too recent formulation of the resource and capability based
theorizing (e.g. (Barney 1991; Teece, Pisano, and Shuen 1997)). However, the dominance of
the firm-level explanations have been partly refuted by later scholars finding the sum of
indirect and direct effects from industry to be of more importance than previously
indicated by (Mcgahan etc). Moreover, the industry effect is also more time invariant
(i.e. more stable) compared to firm level effects. Recent findings also suggest that there
are considerable interaction effects between industry and the firm indicating that
industry membership works directly at performance as well as indirectly through affecting
the firm. However, little or no explanations are made for the mechanisms of this
interaction.In other words, industry factors matter alas less than firm-level factors, and
the indirect link from industry-level through firms are in need of further development.

>>>>

\documentclass[review,fleqn]{elsarticle}
\usepackage[textsize=tiny,colorinlistoftodos]{todonotes}
%\usepackage{subfigure}
%\usepackage{subcaption}
%\usepackage{caption}
%\usepackage{tikz}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{filecontents}
\usepackage{tikz-qtree}
\usepackage{pdflscape}
\usepackage{tikz}
\usepackage{multirow}
\usepackage{booktabs}
\usepackage{color, colortbl}

  \definecolor{Gray}{gray}{0.9}

\captionsetup[subfigure]{labelfont=bf,textfont=normalfont,singlelinecheck=off,justification=raggedright}
\input{preamble.tex}
%%\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\setlength{\mathindent}{0pt} %% NB MAYBEREMOVE BEFORE SUBMISSION

@@

%\address[mymainaddress]{1600 John F Kfennedy Boulevard, Philadelphia}
%\address[mysecondaryaddress]{360 Park Avenue South, New York}

@@

<<load, results="asis", echo=FALSE >>=

setwd("/Users/larshovdanmolden/Documents/git/article4")
library(xtable)
#load("./data/resart4.RData")
r <- main$resmat
rm  <- main$resmatmed
l  <- main$loadings

foolars=0.99

@
%% Extant literature on dynamic capabilities is battling with a fundamental conception and
%% prediction of the theory, namely if, when and how dynamic capabilities can lead to competitive
%% advantage.
%% This paper is about how apsobritve capacity can come about as strategic posture
%https://arrow.dit.ie/cgi/viewcontent.cgi?article=1009&context=buschmancon

@@

The conventional mechanism explored in the extant literature is related to changes in the
firm`s resource base \citep{Eriksson2014,Schilke2018}. This mechanism is also included in
the dominating definition of dynamic capabilities found in \citep{Helfat2007}: 'A dynamic
capability is the capacity of an organization to purposefully create, extend or modify its
resource base' (p. 4). And although this definition opens up for a variety of components
included in the term 'resource base', a point we discuss later in this paper, most of the
subsequent contributions have focused on changes in operational routines (ref),
capabilities (ref), practices (ref) or tangible assets (ref). This focus, albeit very
important to the development of dynamic capabilities literature, seems to leave out a
considerable part of a firm`s 'resource base', namely those resources that are tied to how
the employees of the firm sees their mission, internalize information and effectuates
change. These traits of the organization is very much related to concepts of
'organizational culture' (ref), 'employee cognition' (ref)' 'organizational genes'
\citep{Nelson1982} and 'sense-making' \citep{Weick1995} to name but a few. However, in
relations with dynamic capabilities one promising avenue for exploring mechanisms beyond
the conventional ones described above, makes a clear distinction between \emph{behavioral}
and \emph{non-behavioral} objects of change \citep{Verona2011}. These non-behavioral
objects of change contain psychological objects such as cognitive frames, motivation and
identity that are important for strategy formation and implementation (Gavetti etc, vince
and gabriel ch 15 of book by verona) as they 'concerns how decisions or actions are
shaped, subverted and transformed by emotions, and (..) determine the "way we do things
here"' (Vince and gabriel). In this respect they work as firm resources in generating
outcomes such as performance and competitive advantage. This notion is perhaps expressed
most clearly by Peter Drucker`s famous quote about how 'culture eats strategy for lunch'
(ref) insofar as. Moreover, these non-behavioral objects of change are themselves far
from static and are found to change over time as responses to external change
\citep{Zollo2016} thus making them a natural part of the mechanisms underpinning how
dynamic capabilities work \citep{Schilke2018,Verona2011}. Still, far less is know about
the non-behavioral objects of change and how they work as a mechanism in the dynamic
capabilities theory.

@@

To shed light on this important question we theorize on the role of a particular type of
non-behavioral object of change, \emph{strategic cognition}, and how it acts as a
mediator between dynamic capabilities and firm competitive advantage and hence has a
pivotal place in the literature. We build on important work on the concept of
organizational cognition (ref), non-behavioral object of change (ref) and cognitive
psychology (kahneman and tversky),
and hypothesize about how strategic cognition acts as a mediator in the same way as
conventional behavioral objects of change (such as routines and capabilities) and that
these two are complementary to a certain degree. Consequently, we argue, dynamic capabilities takes
the function of a 'conductor' by orchestrating routines and capabilities, and the role of
a 'coach' in stimulating strategic cognition. We use the case of Microsoft's transition
into the cloud computing business as an illuminating case for demonstrating how the roles
of 'conductor' and 'coach' coexist and act partly complimentary in generating strategic
change and competitive advantage.

SOMETHING ABOUT COMP ADV / why interesting outcome holy grail helfat schilke

@@

\section{Definitions}
The collective term for assets available for firms in their pursuit of their objectives
are \emph{resources}. This term entails tangible assets such as machines, capital, and
labor, but also intangible assets such as procedures, practices, and intellectual
property. The process of using these resources in the pursuit of business objectives such
as competitive advantage is a firm's \emph{capabilities} which can be a set of routines
and skills applied. Resources and capabilities are both tied organizational behavior and
how the organization act in their operation. Hence, they are coined \emph{behavioral
  objects of change} \citep{Verona2011}: 'A pattern of action, a process, an operating
routine, or any form of group activity characterized by some level of stability and
predictability' (ibid. p.538). Such behavioral objects such as capabilities are the
conventional understanding of how dynamic capabilities work; they 'purposefully create,
extend and modify the resource base' (helfat). In the terminology proposed in this paper
this process takes the shape as a 'conductor'.

However, as \cite{Verona2011} points out: 'This approach misses the fundamental aspect
that organizational capabilities specific to managing change can hardly be reduced to the
management of behavioral change' (ibid p. 538). This opens for \emph{non-behavioral
  objects of change} which are defined as 'behavioral antecedents, such as cognition
(e.g. cognitive frames) and motivation' (ibid p. 539). This concept entails the 'how
decisions or actions are shaped, subverted, and/or transformed by emotions; and it
concerns how emotions become embedded in cultural and political practices that determine
the "way we do things here"' (Vince and Gabriel 2011 . ch 15). One particular dimension of
these non-behavioral objects of change particularly conducive to the context of
generating strategic change is the \emph{strategic cognition} of the
organization. Strategic cognition can be defined as conceptual structures in the minds of
the individuals that encapsulate a shared understanding of the reality the
individuals face with respect to the strategic goal of the organization \footnote{This
  definition is based on the definition of \cite{Lakoff1987} definition of the more generic
  'cognitive representation' and is adapted to fit the pursuit of strategic goals such as
  competetive advantage which is the aim of this paper}. This strategic
cognition is the result of a mental process that underlie internal
audicences adoption of new strategic representation and identity code to move toward a
new or existing goal \citep{Gavetti2012}. Although strategic cognition may lead to
behavioral outcome, the nature of the concept itself is tied to cognition rather than
actual latent action \citep{Zollo2016}. Thus, dynamic capabilities are not sufficiently
creating strategic change through its role as 'conductor'. Additionally they need to
exercise a role of 'coach' to adapt the strategic cognition of the organization to create
a shared understanding of the sense and necessity of the strategic goals of the
orientation.

@@

%DC as resource orchestration - influencing capabilitie
%DC as cognitive programmer - inluencing cognitive frames

%Prior research focused on managerial cognition, but gavetti shows that hierarchy matters
%and different cognition maters

@@

Satya Nadella is very much on the same page as recent contributions on how cognitive
frames of an organization (i.e. its non-behavioral object of action) is a pivotal part in
how firm resources shape outcomes such as performance and competitive advantage
\citep{Barney1991a,Barney2001}, and how it consequently plays a role in how dynamic
capabilities work along these lines \cite{Verona2011}. Microsoft under Nadella did
transform the company by changing the behavioral objects of action (i.e. capabilities and
routines), but also through changing the non-behavioral objects of action (i.e. the
strategic cognition). In other words, the story of Microsoft`s transformation suggest that
dynamic capabilities worked through mediating factors to influence competetive advantage,
but that these mediating factors exist as two distinctly different concepts working in
tandem. Figure \ref{fig:fig1} presents a conceptual depiction of the paths running through
these two concepts of \emph{capabilities} and \emph{strategic cognition} objects of action
\citep{Zollo2016}. We use a familiar term for the conventional understanding of dynamic
capabilities mechanism as 'orchestration' (ref) as a collective term to describe the
behavioral objects of action (ref). In this process dynamic capabilities play the role of
a 'conductor' orchestrating capabilities of the firm, an understanding that is in line
with a range of the most common definitions of dynamic capabilities \citep{Schilke2018}.

The non-behavioral objects of action has no similar agreed upon analogy, so for the
purpose of our analysis we envision this mechanism to work more as a cognitive process
enabling the organization to tear away from their current understanding and ways in which
things are done (Vince and Gabriel). Whereas the analogy of a 'conductor' suggests a
function that composes an outcome based on the available capabilities, the notion of a
'coach' refers to the function of enabling others to achieve their objectives. The process
is all about changing an organization`s thinking from the automated and well-known, to the
more cognitively aware, moving from what Kahneman and Tversky coins the automated 'system
one' to the deliberate 'system two' thinking (ref). We will now turn to the two functions
of dynamic capabilities in turn.

@@

\subsection{Capabilities conductor}

Satya Nadella set out to transform Microsoft by investing heavily in artificial
intelligence and cloud computing capabilities. This entailed hiring new people, upgrading
the skills and abilities of the current staff, and acquire infrastructure and
hardware. Additionally, Microsoft was able to acquire and integrate a number of companies
to support the strategic direction the company was on. Companies such as GitHub, LinkedIn
and Citrus data became important platforms for developing new capabilities within the
company (Nadella).

This role of 'conductor' is the conventional way of understanding dynamic capabilities. It
stems from one of the most common definitions \citep{Helfat2007} as well as the analogy of
the process as 'resource orchestration' \citep{Sirmon2011}. It is also found in the
empirical literature handling various functional forms of dynamic capabilities such as
acquisition targeting \citep{Bingham2015}, alliancing \cite{Schilke2014}, new product
development \cite{Danneels2008}, innovation capabilities \citep{Breznik2014}, marketing
capabilities \cite{Brune2009} and alliance capabilities \cite{Kale2002,kale2007}.

The bulk of prior contributions study firm capabilities and resources as the mechanism
through which dynamic capabilities influence an outcome
\citep{Fainshmidt2016a,Protogerou2012,Schilke2018,Eriksson2014}. Particularly, much of the
extant literature concerns measures of accounting profitability and competitive advantage
as the final outcome of dynamic capabilities
\cite{shamsie2009,Schilke2014,Schilke2014a,Teece2016}. Thus, from these prior
contributions we derive a fundamental hypotheses of the relationship between dynamic
capabilities and competitive advantage:

\emph{H1: Firm capabilities mediates the relationship between dynamic capabilities and
  competetive advantage}

However, this mechanism is solely related to behavioral objects of action while previous
insights from the literature on organizational capabilities hold that this is only parts
of the story \cite{Tripsas2000,Gavetti2012}. Or as highlighted by \cite{Verona2011}: Most
'conceptualizations of DCs have in common the often implicit assumption that the object
upon which DCs produce their effects is fundamentally of a behavioral nature: that is a
pattern of action, a process, an operating routine, or any form of group activity
characterized by some level of stability and predictability. This approach misses the
fundamental aspect that organizational capabilities specific to managing change can
hardly be reduced to the management of \emph{behavioral} change' (p. 538). Thus, they
argue, a more holistic approach including cognition is needed \citep{Zollo2016}. This claim is also
supported in the lessons from Microsoft's transition.

% https://hbswk.hbs.edu/item/the-transformation-of-microsoft

\subsection{Cognitive coach}

On February 4th 2014 an email appeared in the inbox of all Microsoft employees. It was
from Nadella who had just begun his tenure as CEO of the company. In this email he lays
out the framework for transforming Microsoft around changing the way the company jointly
acted. He wanted to change the culture of the company from one of inviting new ideas and
fostering creativity. This called for changes in the \emph{cognition} of the individual
employers as well as their capabilities and skills. Nadella made this abundantly clear in
this email even paraphrasing Oscar Wilde in that 'we need to believe in the impossible and
remove the improbable'. Nadella's tool for achieving this change of cognition was to
become the notion of the growth mindset \citep{Dweck2016}. This transformation of the
cognition of the organization to better fit the strategy he was laying out, is a proper
example of instilling a \emph{strategic cognition} in the organization.

Juxtaposed to the experiences from Microsoft, the case of Polaroid brings another set of
experiences to our discussion on non-behavioral objects of action (see \citep{Tripsas2000}
for a detailed account). Cognitive inertia on the part of the organization played a
pivotal part in the failing of the company`s digital imaging effort
\citep{Verona2011}. In other words, Polaroid did not exercise dynamic capabilities to
change the cognition of the organization. Similar patterns are found in Kodak and Anderson Consulting (ref
Henderson and Kaplan 2005), and Smith Corona (Danneels 2010). These examples highlight the
absence of 'dynamic capabilities specific to the adaptation of cognitive frames, first,
and consequently of operating processes' \citep[p. 541]{Verona2011}. This is also
highlighted at the core of the dynamic capabilities literature: 'Dynamic capabilities are about doing the
right things, at the right time, based on new product (and process) development, unique
managerial orchestration processes, a strong and {\bf change-oriented organizational
  culture}' (our emphasis) \cite[p]{Teece2014}.

When dynamic capabilities are executed it is as a response to changes in the environment
in order to stay competitive. This processes relates to creating strategic change
\citep{Helfat2007} and entails changing strategic direction and objectives of the
organization. This strategic change requires, in addition to modifying, extending and
creating firm capabilities, also some kind of alignment with the organization as a whole -
what is often referred to as 'how we do things here' \citep{Nelson1982}, and how an
organization thinks and act (NAdella). This alignment is possible only if the employees
share a common cognitive frame related to the the strategic direction of the company
enabling them to make sense of the changes ahead. This is what we have coined
\emph{strategic cognition} and has its expression in Microsoft's development of a growth
mindset. Strategic cognition is needed to get employees on board as well as capture the
efficiency of non-directed change. Getting on board refers to the general willingness of
the organization to accept and value the importance of some decision or system 'without
being prompted or required by formal governance
mechanisms'\citep[p. 812]{Simsek2009}. Moreover, When individuals of a firm buys into
changes they will enhance the full organization`s ability to not only act upon changes
(Marchand et.al 2000), but also in 'timely responsiveness' (Kohli et.al 1993,
Simsek2009). Consequently, strategic cognition not only contributes to getting people on
board, it also maintain a more efficient change process where the individuals take more
responsibilities for the change itself and leverages their individual knowledge in the
endeavor. For a firm to create real strategic change as enabled by their dynamic
capabilities and hence stay competitive, is will need to act as a 'coach' to instill a
strategic cognition in the organization.

\emph{H2: Strategic cognition mediates the relationship between dynamic capabilities and
  competitive advantage}

\subsection{Proactiveness as strategic cognition}
